---
title: "00 Seed Germination Retesting List"
author: "Katie Heineman"
date: "6/4/2019"
output: html_document
---

# How to create a procedure for retesting, while still gaining data for publication?

# Assumptions

+ Every quarter we can run 30 tests (120 test per year)
+ Each year we have around 20 rare seed collections

# Data Import
```{r set up, message=FALSE, warning=FALSE, paged.print=FALSE}
library(data.table)
library(httr)
library(RCurl)
library(rjson)
library(jsonlite)
library(DT)

# Reading in Caspio Token
source("Tokens/caspioSource.R")
# Reading in caspio GET command (don't know if this works for things > 1000)
source("caspioFunctions.R")

# Read in Datasets from Caspio API where possible, but also from static sources where necessary
germTest <- caspio_get_table("Germination",login1)
seedAcc <- caspio_get_view("Seed_Accession_All_Info",login1)
germLog <- fread("Data/GerminationLog_2019-Jun-04_2151.csv")

# Select Columns from desired datasets
seedAccLim <- seedAcc[Deaccessioned==0,.(SeedAccessionNumber,ScientificName,CNPSRank,LocalityName,DateSeedCollected,MaternalLinesSampled,DateFrozen)]
```

# Make Table of Test Status for Each Accession

```{r status1, message=FALSE, warning=FALSE}

## Finding inital counts of germinants
# Summarize Germination Test Initial Counts for those with initial values in the initital count columns (new database format)
germTestInitLong <- melt(germTest[!is.na(Rep1Init)],
                     id.vars=c("PK_ID","SeedAccessionNumber"),
                     measure.vars = names(germTest)[grepl("Init",names(germTest))]
                     ,variable.name="Rep"
                    , value.name="InitialCount",na.rm=T) # this code is not performing correctly
# Remove "_Init" from rep column
germTestInitLong[,Rep:=gsub('Init', '', Rep)]

# Summarize Germination Test Initial Counts & # of Germinants from test in the old format
germTestInitOld <- germTest[is.na(Rep1Init) & !is.na(NumberSeedsPulled),.(PK_ID,SeedAccessionNumber, InitialCount=NumberSeedsPulled,NumGerminantsOld=NumGerminants)]
germTestInitOld[,Rep:="Old"]

# Bind two initial tests together
germTestLong <- rbind(germTestInitLong,germTestInitOld,fill=T)


## Reshaping Germination Log for New tests
# melt germLog so that we have one observation for each rep
germLogMelt <- melt(germLog, id.vars = c("GermLogID","GermTestID","DateChange", names(germLog)[15:18]), measure.vars = names(germLog)[5:14], variable.name="Rep",value.name="measurementCount",na.rm=T)

# Remove "_Count" from rep column
germLogMelt[,Rep:=gsub('_.*', '', Rep)]

# Merge together log and germ tests
germMerge <- merge(germTestLong, germLogMelt, by.x=c("PK_ID","Rep"), by.y=c("GermTestID","Rep"), all.x=T)

## Summing Germination for Both Old & New Test Types
# Count the Germinants Per Rep - Need to fix the creation of Infinity
germCountsbyRep = germMerge[,.(InitialCount=(InitialCount[1]), MaxCount = max(as.numeric(measurementCount),na.rm=T),NumGerminantsOld=NumGerminantsOld[1]), by=c("SeedAccessionNumber","PK_ID", "Rep")] 
# Summarizing Germinants by Test (taking into account old tests)
germCountsbyTest  = germCountsbyRep[,.(SeedsTested=sum(InitialCount), GerminantsNew =sum(MaxCount,na.rm=T),NumGerminantsOld=NumGerminantsOld[1] ), by=c("SeedAccessionNumber","PK_ID")]
# Calculate Percent Germination
germCountsbyTest[,Germinants:=ifelse(GerminantsNew==-Inf, NumGerminantsOld,GerminantsNew)]
germCountsbyTest[,GermPercent:=round((Germinants)/(SeedsTested)*100)]
germCountsbyTest[,NumGerminantsOld:=NULL]
germCountsbyTest[,GerminantsNew:=NULL]

## Merge back on important fields
germCountsbyTest <- merge(germCountsbyTest, germTest[,.(PK_ID, StartDate,FinalDate,TestType, Baseline)], by="PK_ID")

## Make a Table of the Test Types Completed
germinationSummary <- germCountsbyTest[,.(BaselineYN=ifelse(grepl("Initial",toString(TestType)),"Yes","No"), 
                                       BaselineStatus=toString(unique(Baseline[TestType=="Initial"])),
                                       BaselinePercent=toString(GermPercent[TestType=="Initial"]),
                                       FollowUpYN=ifelse(grepl("Follow-up",toString(TestType)),"Yes","No"),
                                       FollowUpPercent=toString(GermPercent[TestType=="Follow-up"])),by="SeedAccessionNumber"]

## Add this table to overall accession table
overallSummary <- merge(seedAccLim,germinationSummary,by="SeedAccessionNumber",all.x=T)

## Fixing dates
overallSummary[,DateSeedCollected:=as.Date(gsub("T00:00:00","",DateSeedCollected))]
overallSummary[,DateFrozen:=as.Date(gsub("T00:00:00","",DateFrozen))]

## Change aspects to allow sorting
overallSummary[is.na(CNPSRank),CNPSRank:="No Rank"]
overallSummary[is.na(BaselineYN),BaselineYN:="No"]

## Converting CNPS so that it is easier to sort 
overallSummary[grepl("1B",CNPSRank),CNPSRank:="1B"]
overallSummary[grepl("2B",CNPSRank),CNPSRank:="2B"]
overallSummary[grepl("3",CNPSRank),CNPSRank:="3"]
overallSummary[grepl("4",CNPSRank),CNPSRank:="3"]


datatable(overallSummary[,.(Acc=SeedAccessionNumber,Spp=ScientificName,CNPSRank, Local=LocalityName, DateFrozen,BaselineYN,BaselineType=BaselineStatus,FollowUpYN)])
```

# Priorities for Fresh Testing Summer 2019

I think the fastest way to get good retesting is to prioritize fresh testing of **rare** accessions that were frozen in 2018 that are due for retesting in 2018. Ideally, we do first year retests of the accession frozen in 2018 and then slowly work our way backwards to test the rest of the rare accessions over time? What do you think?

Is there a particular date before which we have no fresh seed? I am assuming that we have no fresh seed from before 2015 here. I am prioritizing accessions by CNPS rank and then by collection date. In total, there are 64 recent accessions with no fresh test

```{r fresh priorities, message=FALSE, warning=FALSE}
# Prioritizing 2018 accessions first
overallSummary[,PrioritySort:=ifelse(year(DateFrozen)=="2018",1,2)]
overallSummary[is.na(PrioritySort),PrioritySort:=ifelse( year(DateSeedCollected)=="2018",1.5,2)]
overallSummary[is.na(PrioritySort),PrioritySort:=2]

setorder(overallSummary,PrioritySort,DateFrozen,CNPSRank,DateSeedCollected)

priorityTableFresh <- overallSummary[BaselineYN=="No" & DateSeedCollected>"2015-01-01" & CNPSRank!="No Rank",.(Acc=SeedAccessionNumber,Spp=ScientificName,CNPSRank,LocalityName,Collected=DateSeedCollected, Frozen=DateFrozen)]

# Write out Priority Table
currentDate <- Sys.Date() 
write.csv(priorityTableFresh,paste0("Data/priorityTableFresh-", currentDate,".csv"))

# Printing Table for View
datatable(priorityTableFresh)

```

# Frozen Priorities for Fall 2019

We want to start by retesting the rare accessions due for a 1 year retest in 2019. This list is similar to the fresh priority list except that it obviously includes accessions that already have baseline tests. 

```{r status, message=FALSE, warning=FALSE}

priorityTableFrozen <- overallSummary[CNPSRank!="No Rank",.(Acc=SeedAccessionNumber,Spp=ScientificName,CNPSRank,LocalityName,BaselineYN,Collected=DateSeedCollected,PrioritySort, Frozen=DateFrozen)]

# Write out Priority Table
currentDate <- Sys.Date() 
write.csv(priorityTableFrozen,paste0("Data/priorityTableFrozen-", currentDate,".csv"))

# Printing Table for View
datatable(priorityTableFrozen)

```
